---
title: 'ggplot2 intro exercise'
author: "Mike Frank"
date: "November 16, 2022"
output: html_document
---
  
This is an in-class exercise on using `ggplot2`.

Note, that this example is from the_grammar.R on [http://had.co.nz/ggplot2](). I've adapted this for psych 251 purposes.

First install and load the package. It's part of the "core tidvyerse".

```{r}
library(tidyverse)
```

We're going to be using the `diamonds` dataset. This is a set of measurements of diamonds, along with their price etc.

```{r}
head(diamonds)
```

# Simple ggplots

Remember, the three minimal things you need for a ggplot are:

1. a `ggplot()` call with the data frame as your first argument. 
2. an aesthetic `aes()` statement within that ggplot call to set up the  mappings of visual variables to variables in the data frame.
3. a geometric layer to turn that mapping into marks on the plot, e.g. `geom_point(). 

First let's try making a scatter plot of diamonds' carats  plotted by their price.

```{r}
ggplot(diamonds, aes(x = carat, y = price)) + 
  geom_point()
```

Now visualize the `clarity` variable, mapping it onto point color. Hint: add color to the aesthetic. 

```{r}
...
```



# Building more complex plots

ggplot allows you to use the `+` operator to add objects together into a plot. Then calling that plot (here `p`) actually renders the plot.

```{r}
p <- ggplot(diamonds, aes(x=carat, y=price)) # first you set the aesthetic and dataset

p <- p + geom_point(aes(color = clarity)) # and you can  add layers to the plot

p
```

Try writing this same plot as a single set of additions (e.g. one line of R code, though you can put in linebreaks). 

```{r}
...
```

Now let's try adding some extra bits to the plot. First we'll add a smoothing line. Take the plot above and add a `geom_smooth` layer.  

```{r}
...
```

We can also make use of transparency to make the points more readable. Try adding something to this same plot by changing the `alpha` value for the points by adding `alpha=.02` to your `geom_point` call. Try some different alpha values (bigger and smaller) to see what they do.

```{r}
...
```

See how this makes your plot look different?

# Facets

One of the primary benefits of `ggplot2` is the use of facets - also known as "small multiples." 

Let's take your plot above, but let's break it up into many small multiples based on the clarity. 

We can do this with the command `facet_wrap()`. You can write `facet_wrap(vars(x))` if you want to create facets by breaking down the data on variable `x`. 


```{r}
...
```

You can also use `facet_grid(vars(x, y))` to amke a grid of small facets based on variables `x` and `y`. Try doing this for `color` and `clarity`.


```{r}
...
```


# Geoms

As you've seen above, the basic unit of a ggplot plot is a "geom" - a mapping between data (via an "aesthetic") and a particular geometric configuration on coordinate axes. 

Let's try working with some summary data now. 

```{r}
diamonds_summary <- diamonds |>
  mutate(carat_bin = ceiling(carat * 4)/4) |> # bin by .25
  group_by(carat_bin, cut, color, clarity) |>
  summarise(price = mean(price),
            n = n(), 
            ci = 1.96 * sd(price) / sqrt(n()))
```

Now let's plot the summary for each bin.

```{r}
ggplot(diamonds_summary, aes(x = carat_bin, y = price)) + 
  geom_point() 
```

Let's add some structure to this plot: use visual variable mappings and facets to show some or all of the variables `cut`, `clarity`, and `price`. 

```{r}
...
```

CHALLENGE PROBLEM 1: can you plot error bars on each summary? I've given you a `ci` for each obseration in the data frame. 

```{r}
...
```


CHALLENGE PROBLEM 2: can you show the amount of data in each bin visually using the `n` variable?

```{r}
...
```


# Themes and plot cleanup

I like a slightly cleaner look to my plots. Luckily, ggplot allows you to add "themes" to your plots. Try doing the same plot but adding `+ theme_bw()` or `+ theme_classic()`. Different themes work better for different applications, in my experience. 

You can also try different color scales. Experiment with `scale_color_...` - try writing that and hitting TAB for autocomplete options. Check out the help on this. 

You can also try transparency/different point sizes to clean up scatterplots. Try `alpha = .1` or `pch = "."` to make transparent or small points. 

Finally, don't forget to "fix the axis labels"! 

Here's a "fine" (ugly) plot - see if you can make it look awesome. 

```{r}
ggplot(diamonds, aes(x = carat, y = price, 
                     col = cut)) + 
  geom_point() + 
  facet_wrap(~clarity) 
```

Now you try it:

```{r}


```
